---
title: "A linear regression model"
author: "Rong Zhou"
date: "1/22/2022"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This example is taken from 
A Modern Approach to Regression with R
https://gattonweb.uky.edu/sheather/book/

# Set up the work space

First I want to reduce the number of displayed digits.

```{r}
options(
  digits = 4
)
```

# 1. Load the data

In R, there is a special function for reading a data set into R. For .txt files, you will probably use read.table().

Use file.path(choose.files()) to get a data set's file path in Windows.

```{r}
production <- read.table(
  file = "C:\\zrwork\\UCF\\2022Spring\\4164\\Files\\Production.txt",
  header = TRUE
)
```

Let's look at our data.

```{r}
production
```

#2. Look at the dataset

The head() and tail() functions print the first and last rows of a dataset. This is an easy way to quickly get a feel for your data.

It is not possible to visually inspect large datasets. We work with large datasets by taking snapshots and summaries of the data.

Let's print the dataset and limit our print to include only the first ten observations.

```{r}
head(production,10)
```

Let's print the dataset and limit our print to include only the last ten observations.

```{r}
tail(production,10)
```

#3. Summarise the data

The summary() function prints summary statistics of a dataset.

```{r}
summary(
  object = production
)
```

# Size of the data

When processing data, you need to know how much data you started with and how much you currently have. The dim() function returns the number of rows and columns; nrow() returns the number of rows. ncol() returns the number of columns.

complete.cases() returns a logical vector indicating which rows have zero missing values. 

```{r}
dim(production)
nrow(production)
ncol(production)
sum(complete.cases(production))
```

This is a regression class, so let's plot a scatter plot. We do this to see if building a regression model is reasonable.

```{r}
par(
  mfrow=c(1,1)
)
plot(
  x = production$RunSize,
  y = production$RunTime,
  xlab="Run Size", 
  ylab="Run Time"
)
```

# A simple linear regression model

Let's build a simple linear regression model.

```{r}
lm_production <- lm(
  formula = RunTime~RunSize,
  data = production
)
```

Now let's put the linear model on our scatter plot to see if the model is reasonable.

```{r}
plot(
  x = production$RunSize,
  y = production$RunTime,
  xlab="Run Size", 
  ylab="Run Time"
)
abline(
  reg = lm_production,
  col = "red"
)
```

# Hypothesis testing

Hypothesis testing is critical to small data analytics. anova() and summary() give us the most important tests for linear regression.

```{r}
anova(
  object = lm_production
)
summary(
  object = lm_production
)
```

Let's check those t-values from the summary() output.

```{r}
149.7477/8.3282
0.2592/0.0371
```

# Confidence intervals for the coefficients

Now let's look at confidence intervals for the coefficients.

```{r}
confint(
  object = lm_production,
  level=0.95
)
```

# Expected values and predictions

```{r}
score_data <- data.frame(
  RunSize = seq(
    from = min(production$RunSize),
    to = max(production$RunSize),
    by = 5
  )
)
head(score_data)

predict_none <- predict(
  object = lm_production,
  newdata = score_data,
  interval = "none",
  level=0.95
)
head(predict_none)

predict_confidence <- predict(
  object = lm_production,
  newdata = score_data,
  interval = "confidence",
  level=0.95
)
head(predict_confidence)

predict_prediction <- predict(
  object = lm_production,
  newdata = score_data,
  interval = "prediction",
  level=0.95
)
head(predict_prediction)
```

Now let's add the 
expected values/predictions (red),
expected value confidence intervals (blue),
prediction intervals (purple).

```{r}
plot(
  x = production$RunSize,
  y = production$RunTime,
  xlab="Run Size", 
  ylab="Run Time"
)
abline(
  reg = lm_production,
  col = "red"
)
lines(
  x = score_data$RunSize,
  y = predict_confidence[,"lwr"],
  col = "blue"
)
lines(
  x = score_data$RunSize,
  y = predict_confidence[,"upr"],
  col = "blue"
)
lines(
  x = score_data$RunSize,
  y = predict_prediction[,"lwr"],
  col = "purple"
)
lines(
  x = score_data$RunSize,
  y = predict_prediction[,"upr"],
  col = "purple"
)
```
Save the confidence intervals into a excel file:

```{r}
write.csv(predict_confidence,file = "C:\\zrwork\\UCF\\2022Spring\\4164\\Files\\confidence_interval.csv")

write.csv(predict_prediction,file = "C:\\zrwork\\UCF\\2022Spring\\4164\\Files\\prediction_interval.csv")
```
